---
title: "ESM 204 HW 4: Calculating the SCC and policy choice under uncertainty"
author: "Larissa Neilson, Alison Sells, Katelyn Toigo"
date: "5/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(ggpubr)
library(patchwork)
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Read in the data
damage <- read_csv(here("data", "damages.csv")) %>% 
  clean_names()
warming <- read_csv(here("data", "warming.csv")) %>% 
  clean_names()
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Question 1
damages <- damage %>%
  mutate(warming2 = warming^2)

# Quadratic model of damages
dam_lm <- lm(damages ~ warming + warming2, data = damages)
dam_lm[["coefficients"]][["(Intercept)"]] <- 0

# Damages Plot
damages_plot <- ggplot(data = damages) +
  geom_point(aes(x = warming, y = damages)) +
  stat_smooth(data = damages, aes(x = warming, y = damages)) +
  labs(x = "Level of warming (ºC)", y = "Annual total damages ($)") +
  theme_minimal()
#damages_plot
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Function with our quadratic equation
damage_function <- function(warming) {
  damages <- (19500000000000 * warming^2) - (3019000000000 * warming)
  return(damages)
}

# Test
#damage_function(0.3022845)
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Test, this should input the same # as the last test line but it's slightly off?
#damage_function(warming[1, "warming_baseline"])

# This makes new column with the calculated damages using values from the warming_baseline column
warming$damage_base <- damage_function(warming$warming_baseline)
#head(warming$damage_base)

# Question 2 plot (1)
plot_baseline <- ggplot(data = warming, aes(x = year, y = damage_base)) +
  geom_point(cex = 0.5) +
  labs(x = "Year", y = "Total damages ($)", title = "Damages without Pulse") +
  theme_minimal()

# Do the same with values from the warming_pulse column
warming$damage_pulse <- damage_function(warming$warming_pulse)
#head(warming$damage_pulse)

# Question 2 plot (2)
plot_pulse <- ggplot(data = warming, aes(x = year, y = damage_pulse)) +
  geom_point(cex = 0.5) +
  labs(x = "Year", y = "Total damages ($)", title = "Damages with Pulse") +
  theme_minimal()
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
warming_diff <- warming %>% 
  mutate(damage_diff = (damage_pulse - damage_base))

# Question 2 plot (3)
plot_difference <- ggplot(data = warming_diff, aes(x = year, y = damage_diff)) +
  geom_point(cex = 0.5) +
  labs(x = "Year", y = "Difference in damage ($)", title = "Diff. in Damages with Pulse") +
  theme_minimal()
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Question 2 Part 4
warming_diff_norm<- warming_diff %>% 
  mutate(damage_diff_norm = damage_diff/ 35000000000)

# Question 2 plot (4)
plot_difference_per_ton <- ggplot(data = warming_diff_norm, aes(x = year, y = damage_diff_norm)) +
  geom_point(cex = 0.5) +
  labs(x = "Year", y = " Difference in damage ($ per ton)", title = "Diff. in Dmg. with Pulse/ton of CO2") +
  theme_minimal()
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Use patchwork package
all_plots <- (plot_baseline + plot_pulse) / (plot_difference + plot_difference_per_ton)

all_plots
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Question 3
scc <- warming_diff_norm %>% 
  mutate(scc_1 = damage_diff/ (1+0.01)^x1,
         scc_2 = damage_diff/ (1+0.02)^x1,
         scc_3 = damage_diff/ (1+0.03)^x1,
         scc_4 = damage_diff/ (1+0.04)^x1,
         scc_5 = damage_diff/ (1+0.05)^x1,
         scc1_sum = sum(scc_1),
         scc2_sum = sum(scc_2),
         scc3_sum = sum(scc_3),
         scc4_sum = sum(scc_4),
         scc5_sum = sum(scc_5)) 

scc_df <- data.frame(c(0.01, 0.02, 0.03, 0.04, 0.05)) %>%
  cbind(c(scc$scc1_sum[1], scc$scc2_sum[1], scc$scc3_sum[1], scc$scc4_sum[1], scc$scc5_sum[1]))
colnames(scc_df) <- c("rate", "scc_sum")

scc_plot <- ggplot(data = scc_df) +
  geom_point(aes(x = rate, y = scc_sum)) +
  theme_minimal() + ggtitle("SCC vs. Discount Rates")
scc_plot
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Question 4
rams = 0.001 + (2*0.01) # new r value

scc<- warming_diff_norm %>% 
  mutate(scc_rams = damage_diff/ (1 + rams)^x1,
          scc_rams_sum = sum(scc_rams))
         
scc_rams_df <- data.frame(c(0.021)) %>%
  cbind(c(scc$scc_rams_sum[1]))
colnames(scc_rams_df) <- c("rate", "scc_sum")

# This line fixes the legend to be what we want the descriptor to say
cols <- c("SCC Given the Conditions" = "tomato2")

scc_rams_plot <- ggplot() +
  geom_point(data = scc_df,
             aes(x = rate, y = scc_sum)) +
  geom_point(data = scc_rams_df, aes(x = rate, y = scc_sum, color = "SCC Given the Conditions")) +
  theme(legend.title = element_blank()) +
  theme_minimal() + ggtitle("SCC vs. Discount Rates")
scc_rams_plot
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Question 5
warming_sub <- warming %>% 
  select(x1,year, warming_baseline, damage_base) 

prob_dam <- warming_sub %>% 
  mutate(warming_1.5 = warming_baseline * 1.5)

prob_dam$damage_1.5 <- damage_function(prob_dam$warming_1.5)
#head(prob_dam$damage_1.5)

sum_prob_dam <- prob_dam %>% 
  mutate(npv_1.5 = damage_1.5/ (1+0.02)^x1,
         npv_base = damage_base/ (1+0.02)^x1,
         sum_dam_1.5 = sum(npv_1.5), 
         sum_dam_base = sum(npv_base),
         exp_value = ((sum_dam_1.5*0.5)+ (sum_dam_base*0.5)))
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
warming5B <- warming %>% 
  select(x1, year, warming_baseline) %>% 
  mutate(warming = case_when(
    year < 2050 ~ warming_baseline,
    year >= 2050 ~ 1.29),
    dam_b = (warming*3019000000000) + ((warming)^2*19500000000000),
    dam_b_sc = dam_b/ (1+0.02)^x1)

warming_5b_sum <- warming5B %>%  
 mutate(sum_b_sc = sum(dam_b_sc))
```

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Cost of Policy A = Cost of Policy B + x 
# Need to find how much x could be for it still to make sense to use Policy B

# Maximum of x for Policy B to still make sense
exp_value_calc_a <- sum_prob_dam %>% 
  select(exp_value)

exp_value_calc_b <- warming_5b_sum %>% 
  select(sum_b_sc)

max_price_x = (exp_value_calc_a - exp_value_calc_b)
 
#max_price_x
```

If society is risk neutral, Policy B would make economic sense as long as it is less than or equal to $1.979412e+15 over the 80 year time frame. If the cost of implementing Policy B is above $1.979412e+15, then Policy A would make more economic sense.

If society is risk averse, the maximum price of x could increase because the utility of guaranteed temperature reduction in Policy B is more than the utility of Policy A which involves risk.

If society is instead risk adverse, society is willing to pay more for Policy B (up to a certain point) as it eliminates any risk that is present in the business as usual Policy A. The certain point to which a risk averse society would switch to Policy A would be when the price of x would __ the utility of policy of A is greater than the utility of Policy B.

If A and B + x are equal the society will pick Policy B as the utility.

